---
title: 'Project 6: Randomization and Matching'
output: pdf_document
---

# Introduction

In this project, you will explore the question of whether college education causally affects political participation. Specifically, you will use replication data from \href{https://papers.ssrn.com/sol3/papers.cfm?abstract_id=1409483}{Who Matches? Propensity Scores and Bias in the Causal Eﬀects of Education on Participation} by former Berkeley PhD students John Henderson and Sara Chatfield. Their paper is itself a replication study of \href{https://www.jstor.org/stable/10.1017/s0022381608080651}{Reconsidering the Effects of Education on Political Participation} by Cindy Kam and Carl Palmer. In their original 2008 study, Kam and Palmer argue that college education has no effect on later political participation, and use the propensity score matching to show that pre-college political activity drives selection into college and later political participation. Henderson and Chatfield in their 2011 paper argue that the use of the propensity score matching in this context is inappropriate because of the bias that arises from small changes in the choice of variables used to model the propensity score. They use \href{http://sekhon.berkeley.edu/papers/GenMatch.pdf}{genetic matching} (at that point a new method), which uses an approach similar to optimal matching to optimize Mahalanobis distance weights. Even with genetic matching, they find that balance remains elusive however, thus leaving open the question of whether education causes political participation.

You will use these data and debates to investigate the benefits and pitfalls associated with matching methods. Replication code for these papers is available online, but as you'll see, a lot has changed in the last decade or so of data science! Throughout the assignment, use tools we introduced in lab from the \href{https://www.tidyverse.org/}{tidyverse} and the \href{https://cran.r-project.org/web/packages/MatchIt/MatchIt.pdf}{MatchIt} packages. Specifically, try to use dplyr, tidyr, purrr, stringr, and ggplot instead of base R functions. While there are other matching software libraries available, MatchIt tends to be the most up to date and allows for consistent syntax.

# Data

The data is drawn from the \href{https://www.icpsr.umich.edu/web/ICPSR/studies/4023/datadocumentation#}{Youth-Parent Socialization Panel Study} which asked students and parents a variety of questions about their political participation. This survey was conducted in several waves. The first wave was in 1965 and established the baseline pre-treatment covariates. The treatment is whether the student attended college between 1965 and 1973 (the time when the next survey wave was administered). The outcome is an index that calculates the number of political activities the student engaged in after 1965. Specifically, the key variables in this study are:

\begin{itemize}
    \item \textbf{college}: Treatment of whether the student attended college or not. 1 if the student attended college between 1965 and 1973, 0 otherwise.
    \item \textbf{ppnscal}: Outcome variable measuring the number of political activities the student participated in. Additive combination of whether the student voted in 1972 or 1980 (student\_vote), attended a campaign rally or meeting (student\_meeting), wore a campaign button (student\_button), donated money to a campaign (student\_money), communicated with an elected official (student\_communicate), attended a demonstration or protest (student\_demonstrate), was involved with a local community event (student\_community), or some other political participation (student\_other)
\end{itemize}

Otherwise, we also have covariates measured for survey responses to various questions about political attitudes. We have covariates measured for the students in the baseline year, covariates for their parents in the baseline year, and covariates from follow-up surveys. \textbf{Be careful here}. In general, post-treatment covariates will be clear from the name (i.e. student\_1973Married indicates whether the student was married in the 1973 survey). Be mindful that the baseline covariates were all measured in 1965, the treatment occurred between 1965 and 1973, and the outcomes are from 1973 and beyond. We will distribute the Appendix from Henderson and Chatfield that describes the covariates they used, but please reach out with any questions if you have questions about what a particular variable means.

```{r}
# Load tidyverse and MatchIt
# Feel free to load other libraries as you wish
library(tidyverse)
library(MatchIt)
#install.packages("cobalt")
library("cobalt")
# Load ypsps data
ypsps <- read_csv('data/ypsps.csv')
head(ypsps)
```

# Randomization

Matching is usually used in observational studies to to approximate random assignment to treatment. But could it be useful even in randomized studies? To explore the question do the following:

\begin{enumerate}
    \item Generate a vector that randomly assigns each unit to either treatment or control
    \item Choose a baseline covariate (for either the student or parent). A binary covariate is probably best for this exercise.
    \item Visualize the distribution of the covariate by treatment/control condition. Are treatment and control balanced on this covariate?
    \item Simulate the first 3 steps 10,000 times and visualize the distribution of treatment/control balance across the simulations.
\end{enumerate}

```{r}
# Generate a vector that randomly assigns each unit to treatment/control
n = nrow(ypsps)
ypsps <- ypsps %>% 
    mutate(college_random = as.numeric(rbernoulli(n, p=0.5)))

# Choose a baseline covariate (use dplyr for this)
head(ypsps$student_Gen, 10) 
#we selected student Gender 

# Visualize the distribution by treatment/control (ggplot)
ggplot(ypsps, aes(x = student_Gen, fill = factor(college_random))) + 
  geom_bar() +
  facet_grid(college_random~.) + 
  labs(title = "Distribution of Gender among College Attendees vs Non-College Attendees", fill = "college_random\n")

chisq.test(table(ypsps$college_random, ypsps$student_Gen)) 

# Simulate this 10,000 times (monte carlo simulation - see R Refresher for a hint)
gender1 <- list() #empty vector for average treatment given Gender == 1 
gender0 <- list() #empty vector for average treatment given Gender == 0 
iter = 10000
for (i in 1:iter){
    ypsps_sim <- ypsps %>% mutate(college_random = as.numeric(rbernoulli(n, p=0.5)))
    Y0 <- mean(ypsps_sim$college_random | ypsps_sim$student_Gen==0) #return mean of college attendance given Gender==0
    Y1 <- mean(ypsps_sim$college_random | ypsps_sim$student_Gen==1) #returns mean of college attendance given Gender ==1
    gender1 <- c(gender1, Y1) #store into empty vector 
    gender0 <- c(gender0, Y0) #store into empty vector for t
}

#Visualize distribution of gender across college attendance for 10,000 simulations
#unlist
college_gender1 <- as.numeric(unlist(gender1))
college_gender0 <- as.numeric(unlist(gender0))
#save into a dataframe and bind
college_gender1_df <- data.frame(college_gender1)
college_gender0_df <- data.frame(college_gender0)
distribution <- as.data.frame(cbind(college_gender0, college_gender1))


distributions <- distribution %>% 
  pivot_longer(c(`college_gender0`, `college_gender1`), names_to = "collegegender", values_to = "college_attendance")

#plot the distribution
ggplot(distributions, aes(x=college_attendance, fill=collegegender))+ geom_density(alpha=.2) +
  #scale_y_continuous +
  labs(x = "College Attendees", title = "Distribution of Gender among College Attendees vs Non-College Attendees - 10,000 Simulations") +
  labs(scale_fill_discrete(name = "College Attendance", labels = c("Did Not Attend College", "Attended College")))
```


## Questions
\begin{enumerate}
    \item \textbf{What do you see across your simulations? Why does independence of treatment assignment and baseline covariates not guarantee balance of treatment assignment and baseline covariates?}
\end{enumerate}

\textbf{Your Answer}: 
   - We see fairly balanced distribution across genders, i.e. that college attendance was fairly similar between men and women.  
    - There could be an imbalance of treatment assignment and baseline covariates due to just random chance. Also depending on how many baseline covariates you have in your analysis, there might not just be enough people in the combinations of those covariates to balance equally across treatment assignment. In other words, as you increase number of strata (number of covariates to balance across treatment groups), the chance of empty or infrequently filled strata also increases.

# Propensity Score Matching

## One Model
Select covariates that you think best represent the "true" model predicting whether a student chooses to attend college, and estimate a propensity score model to calculate the Average Treatment Effect on the Treated (ATT). Plot the balance of the top 10 (or fewer if you select fewer covariates). Report the balance of the p-scores across both the treatment and control groups, and using a threshold of standardized mean difference of p-score $\leq .1$, report the number of covariates that meet that balance threshold.

```{r}
# Select covariates that represent the "true" model for selection, fit model
ypsps <- ypsps %>% select(-college_random) #remove random Bernoulli from Section 3

postTreat_var_indices <- 123:174 #variables that contain 1983
outcome_var_components <- c("student_vote", "student_meeting", "student_button",
                            "student_money",
                            "student_communicate", "student_demonstrate",
                            "student_community",
                            "student_other")
ypsps_preTreat <- ypsps %>% select(-postTreat_var_indices) %>%  select(-outcome_var_components) 

#LL Note: just went w/ a list of 10 for now, but we can consider expanding it if we'd like!
true_covars <- c('student_PubAff', 'student_Newspaper', 'student_FamTalk', 
                 'student_GovtOpinion', 'student_Govt4All', 
                 'parent_Vote', 'parent_Rally', 'parent_PolClub', 
                 'parent_CivicOrg', 'parent_HHInc')

#true_covars <- c('student_PubAff', 'student_Newspaper', 'student_Radio', 'student_TV', 
#                 'student_Magazine', 'student_FamTalk', 'student_AdultTalk',
#                 "student_FrTalk", 'student_GovtOpinion', 'student_TrGovt',
#                 'student_GovtSmart',
#                 'student_Govt4All', 'student_Govern', 
#                 'student_FDR', 'student_Knowledge', 'student_SchOfficer',
#                 'parent_Newspaper', 'parent_Radio', 'parent_TV', 'parent_Magazine',
#                 'parent_StrOpinion','parent_Vote', 'parent_Rally',
#                 'parent_OthAct', 'parent_PolClub', 'parent_Button', 'parent_Money', 
#                 'parent_GovtOpinion', 'parent_TrGovt', 'parent_GovtSmart',
#                 "parent_Govt4All",
#                 'parent_CivicOrg', 'parent_CLOrg', 'parent_HHInc')

ypsps_true <- ypsps %>% select(true_covars, college, student_ppnscal) %>% drop_na()

#create formula to pass into matchIt function 
ps_formula <- as.formula(
    paste("college", paste(true_covars, collapse = " + "), sep = " ~ ")
)

#nearest-neighbor matching
match_out_NN <- matchit(ps_formula, data = ypsps_true, method = 'nearest', estimand = 'ATT')
NN_summary <- summary(match_out_NN)
NN_summary

#fit data and estimate ATT
match_data_NN <- match.data(match_out_NN)
lm_formula <- as.formula(
    paste("student_ppnscal",  paste('college +', paste(true_covars, collapse = " + ")), 
          sep = " ~ ")
)

fit_NN <- lm(lm_formula, data = match_data_NN, weights = weights)

summary(fit_NN)


# Plot the balance for the top 10 covariates
# <- data.frame(NN_summary$sum.matched)
#NN_df <- tibble::rownames_to_column(NN_df, "covars")
#NN_df <- NN_df %>% mutate(abs_mean_diff = abs(NN_df$Std..Mean.Diff.))  
#arrange(NN_df, abs_mean_diff)

plot(summary(match_out_NN), var.order = 'unmatched', threshold = .1)

# Report the overall balance and the proportion of covariates that meet the balance threshold
```
At least for these ten variables, it appears that the overall balance is worse post-matching on each of the ten covariates chosen. Only one out of the ten covariates meetes the balance threshhold here. 

## Simulations
Henderson/Chatfield argue that an improperly specified propensity score model can actually \textit{increase} the bias of the estimate. To demonstrate this, they simulate 800,000 different propensity score models by choosing different permutations of covariates. To investigate their claim, do the following:

\begin{itemize}
    \item Using as many simulations as is feasible (at least 10,000 should be ok, more is better!), randomly select the number of and the choice of covariates for the propensity score model.
    \item For each run, store the ATT, the proportion of covariates that meet the standardized mean difference $\leq .1$ threshold, and the mean percent improvement in the standardized mean difference. You may also wish to store the entire models in a list and extract the relevant attributes as necessary.
    \item Plot all of the ATTs against all of the balanced covariate proportions. You may randomly sample or use other techniques like transparency if you run into overplotting problems. Alternatively, you may use plots other than scatterplots, so long as you explore the relationship between ATT and the proportion of covariates that meet the balance threshold.
    \item Finally choose 10 random models and plot their covariate balance plots (you may want to use a library like \href{https://cran.r-project.org/web/packages/gridExtra/index.html}{gridExtra} to arrange these)
\end{itemize}

\textbf{Note: There are lots of post-treatment covariates in this dataset (about 50!)! You need to be careful not to include these in the pre-treatment balancing. Many of you are probably used to selecting or dropping columns manually, or positionally. However, you may not always have a convenient arrangement of columns, nor is it fun to type out 50 different column names. Instead see if you can use dplyr 1.0.0 functions to programatically drop post-treatment variables (\href{https://www.tidyverse.org/blog/2020/03/dplyr-1-0-0-select-rename-relocate/}{here} is a useful tutorial).}

```{r message=FALSE}
# Remove treatment and interview ID from the pre-treatment variable dataframe and remove NAs
df_covars <- ypsps_preTreat %>% select(-c("interviewid", 'student_ppnscal', 'college')) %>%  drop_na()

#Remove NAs from pre-treatment variable dataframe
all_data <-ypsps_preTreat %>% drop_na()

set.seed(24)

n_sims <- 5000 #LL note on 3/29: just doing 1000 for now because it always breaks at some point at 10k 
#but haven't been able to debug yet
ATT <- c()
percent_improvement <- c()
beats_threshold <- c()
treatment = 'college'
outcome = "student_ppnscal"

for(sim in 1:n_sims){
    number_of_covars <- sample(1:ncol(df_covars), 1)
    #print(paste("covar number:", number_of_covars))
    vars_index <- sample(1:ncol(df_covars), number_of_covars)
    #print(paste("vars index:", vars_index))
    sample_covars <- df_covars %>% select(vars_index)
    #print(paste("samplecols:", ncol(sample_covars)))
    covar_names <- colnames(sample_covars)

    ps_formula <- as.formula(
        paste(treatment, paste(covar_names, collapse = " + "), sep = " ~ "))

    match_out_NN <- matchit(ps_formula, data = all_data, method = 'nearest', 
                        estimand = 'ATT')
    
    NN_summary <- summary(match_out_NN)
    
    #calculate ATT
    match_data_NN <- match.data(match_out_NN)
    lm_formula <- as.formula(
        paste(outcome,  paste(treatment, paste(covar_names, collapse = " + "), sep = " + "), 
            sep = " ~ "))
    
    fit_NN <- lm(lm_formula, data = match_data_NN, weights = weights)
    lm_NN <- summary(fit_NN)
    ATT_NN <- lm_NN$coefficients[treatment, "Estimate"]

    ATT <- append(ATT, ATT_NN)
    #calculate mean percent improvement
    reduction_df <- as.data.frame.matrix(NN_summary$reduction)
    reduction_mean <- mean(reduction_df$`Std. Mean Diff.`)
    percent_improvement <- append(percent_improvement, reduction_mean)

    #calculate proportion that passed 0.1 threshold
    thresh_count = 0
    matched_df <- as.data.frame.matrix(NN_summary$sum.matched)
    for(val in matched_df$`Std. Mean Diff.`){
        if(abs(val) <= 0.1){
        thresh_count <- thresh_count + 1}
    }
    thresh_prop <- (thresh_count / nrow(matched_df))*100
    
    beats_threshold <-append(beats_threshold, thresh_prop)
        
    }
```


```{r message=FALSE}


ATT <- as.data.frame(ATT)

ATT <- rename(ATT, attnn = "ATT")
ATT <- tibble::rowid_to_column(ATT, "ID")

beats_threshold<- as.data.frame(beats_threshold)
colnames(beats_threshold) <- cbind("thresbt")

beats_threshold <- tibble::rowid_to_column(beats_threshold, "ID")


# Plot ATT v. proportion

mergedatt <- merge(x = ATT, y = beats_threshold, by="ID", all.x=TRUE)

ggplot(mergedatt, aes(x = attnn, y = thresbt)) + 
  geom_point() +
  labs(x = "ATT", y = "Threshold") +
  stat_summary(fun.data= mean_cl_normal) + 
  geom_smooth(method='lm')


# 10 random covariate balance plots (hint try gridExtra)
# Note: ggplot objects are finnicky so ask for help if you're struggling to automatically create them; consider using functions!

#Plot #1
bal.plot(match_out_NN, var.mame="student_GovtOpinion")
```


```{r message=FALSE}
#Plot #2
bal.plot(match_out_NN, var.name="parent_FInc")
```


```{r message=FALSE}
#Plot #3
bal.plot(match_out_NN, var.name="parent_Money")
```


```{r message=FALSE}
#Plot #4

bal.plot(match_out_NN, var.name="student_Cynic")
```


```{r message=FALSE}
#Plot #5
bal.plot(match_out_NN, var.name="parent_InfClub")

```


```{r message=FALSE}
#Plot #6
bal.plot(match_out_NN, var.name="student_TV")

```


```{r message=FALSE}
#plot #7
bal.plot(match_out_NN, var.name="parent_EducW")
```

```{r message=FALSE}
#Plot #8
bal.plot(match_out_NN, var.name="parent_GovtSmart")
```


```{r}
#Plot #9
bal.plot(match_out_NN, var.name="student_GPA")
```

#Plot #10
```{r}
bal.plot(match_out_NN, var.name= "student_FamTalk")
```


## Questions

\begin{enumerate}
    \item \textbf{How many simulations resulted in models with a higher proportion of balanced covariates? Do you have any concerns about this?}
    \item \textbf{Your Answer}:
    \item \textbf{Analyze the distribution of the ATTs. Do you have any concerns about this distribution?}
    \item \textbf{Your Answer:} We are not quite sure what the ATT distribution here is telling us about the threshold, even when we elected to place in a regression line, it is still not clear, at least in this dataset, what the relationship is between the ATT distribution and the covariates that make it above the .1 mean difference cutoff. 
    \item \textbf{Do your 10 randomly chosen covariate balance plots produce similar numbers on the same covariates? Is it a concern if they do not?}
    \item \textbf{Your Answer:} The 10 randomly chosen covariates actually seem a lot more reasonable because they distribibution of the covariate among those who are in the treated vs. the non-treated seems a lot more balanaced. This is even more concerning to us given that the ATT and threshold distribution make even less sense after looking at these 10 random covariates. 
\end{enumerate}

# Matching Algorithm of Your Choice

## Simulate Alternative Model

Henderson/Chatfield propose using genetic matching to learn the best weights for Mahalanobis distance matching. Choose a matching algorithm other than the propensity score (you may use genetic matching if you wish, but it is also fine to use the greedy or optimal algorithms we covered in lab instead). Repeat the same steps as specified in Section 4.2 and answer the following questions:

```{r}
# Remove post-treatment covariates

# Randomly select features
n_sims <- 5000 #LL note on 4/4: just doing 1000 for now, but we can up this and hopefully won't break
ATTam <- c()
percent_improvements <- c()
beats_thresholds <- c()
treatment2 = 'college'
outcome2 = "student_ppnscal"

# Simulate random selection of features 10k+ times

for(sim in 1:n_sims){
    number_of_covars <- sample(1:ncol(df_covars), 1)
    #print(paste("covar number:", number_of_covars))
    vars_index <- sample(1:ncol(df_covars), number_of_covars)
    #print(paste("vars index:", vars_index))
    sample_covars <- df_covars %>% select(vars_index)
    #print(paste("samplecols:", ncol(sample_covars)))
    covar_names <- colnames(sample_covars)

    ps_formula <- as.formula(
        paste(treatment, paste(covar_names, collapse = " + "), sep = " ~ "))

    match_out_mahal_full <- matchit(ps_formula, data = all_data, method = 'full', distance = 'mahalanobis',
                        estimand = 'ATT')
    
    full_summary <- summary(match_out_mahal_full)

# Fit  models and save ATTs, proportion of balanced covariates, and mean percent balance improvement

    #calculate ATT
    match_data_mahal <- match.data(match_out_mahal_full)
    lm_formula <- as.formula(
        paste(outcome,  paste(treatment, paste(covar_names, collapse = " + "), sep = " + "), 
            sep = " ~ "))
    
    fit_mahal <- lm(lm_formula, data = match_data_mahal, weights = weights)
    lm_mahal <- summary(fit_mahal)
    ATT_mahal <- lm_mahal$coefficients[treatment, "Estimate"]

    ATTam <- append(ATTam, ATT_mahal)

    #calculate mean percent improvement
    reduction_df <- as.data.frame.matrix(full_summary$reduction)
    reduction_mean <- mean(reduction_df$`Std. Mean Diff.`)
    percent_improvements <- append(percent_improvement, reduction_mean)

    #calculate proportion that passed 0.1 threshold
    thresh_count = 0
    matched_df2 <- as.data.frame.matrix(full_summary$sum.matched)
    for(val in matched_df$`Std. Mean Diff.`){
        if(abs(val) <= 0.1){
        thresh_count <- thresh_count + 1}
    }
    thresh_props <- (thresh_count / nrow(matched_df2))*100
    
    beats_thresholds <- append(beats_thresholds, thresh_props)
        
    }
```


```{r}
# Plot ATT v. proportion

ATTam <- as.data.frame(ATTam)
colnames(ATTam) <- cbind("attamhl")
ATTam <- tibble::rowid_to_column(ATTam, "ID")


beats_thresholds <- as.data.frame(beats_thresholds)
colnames(beats_thresholds) <- cbind("beats_am")
beats_thresholds <- tibble::rowid_to_column(beats_thresholds, "ID")

# Plot ATT v. proportion

mergedatt2 <- merge(x = ATTam, y = beats_thresholds, by="ID", all.x=TRUE)

ggplot(mergedatt2, aes(x = attamhl, y = beats_am)) + 
  geom_point() +
  labs(x = "ATT", y = "Threshold") +
  stat_summary(fun.data= mean_cl_normal) + 
  geom_smooth(method='lm')

# 10 random covariate balance plots (hint try gridExtra)
# Note: ggplot objects are finnicky so ask for help if you're struggling to automatically create them; consider using functions!


#Plot #1
bal.plot(match_out_NN, var.mame="student_GovtOpinion")
```


```{r message=FALSE}
#Plot #2

bal.plot(match_out_NN, var.name="parent_FInc")
```


```{r message=FALSE}
#Plot #3
bal.plot(match_out_NN, var.name="student_Trust")
```


```{r message=FALSE}
#Plot #4

bal.plot(match_out_NN, var.name="student_TV")
```


```{r message=FALSE}
#Plot #5
bal.plot(match_out_NN, var.name="parent_Govt4All")

```


```{r message=FALSE}
#Plot #6
bal.plot(match_out_NN, var.name="student_TV")

```


```{r message=FALSE}
#plot #7
bal.plot(match_out_NN, var.name="parent_EducW")
```

```{r message=FALSE}
Plot #8
bal.plot(match_out_NN, var.name="parent_GovtSmart")
```


```{r}
#Plot #9
bal.plot(match_out_NN, var.name="student_GPA")
```

#Plot #9
```{r}
bal.plot(match_out_NN, var.name= "student_FamTalk")
```



```{r}
# Visualization for distributions of percent improvement

percent_improvements <- as.data.frame(percent_improvements )
colnames(percent_improvements ) <- cbind("pct_am")
percent_improvements  <- tibble::rowid_to_column(percent_improvements, "ID")


percent_improvements %>% ggplot(aes(pct_am)) +
         geom_histogram(binwidth = 1, fill = "blue", col = "blue") +
         xlab("Percent Improvement Distribution") + 
         ggtitle("Histogram")

```


## Questions

\begin{enumerate}
    \item \textbf{Does your alternative matching method have more runs with higher proportions of balanced covariates?}
    \item \textbf{Your Answer:} Yes, actually! It seems like there's a much clearer relationship between ATT and threshold as well. 
    \item \textbf{Use a visualization to examine the change in the distribution of the percent improvement in balance in propensity score matching vs. the distribution of the percent improvement in balance in your new method. Which did better? Analyze the results in 1-2 sentences.}
    \item \textbf{Your Answer:}
\end{enumerate} Right off the bat, it seems like our new method with full matching and mahalanobis distance calculations did much better. My sense from observing the trends in the data and from discussions in class, matching using nearest neighbors has a huge pitfall insofar as it is really hard to do precise calculations in a multivariate context given that it's only dealing with Euclidean space. This means that there is lots of room for error in making approrpriate unit changes and making sense of distances between incongruous variables. Mahalanobis takes into account the sample's covariance and uses covariance to help standardize the distances between a point P and it's place in the Distribution to which that point P belongs to. Thus, it makes intuitive sense to me why Mahalanobis makes for better estimates given that it produces much better matches relative to the nearest neigbhor approach in our first matching algorithm. 

\textbf{Optional:} Looking ahead to the discussion questions, you may choose to model the propensity score using an algorithm other than logistic regression and perform these simulations again, if you wish to explore the second discussion question further.

# Discussion Questions

\begin{enumerate}
    \item Why might it be a good idea to do matching even if we have a randomized or as-if-random design?
    \item \textbf{Your Answer:} As-if-random designs are still subject to selection bias, thus matching helps with making sure that we are not comparing treatment and control groups that have observed (and ideally unobserved) characteristics. However,, even in randomize design trials, matching might help with validity and increasing of study power. These matches can protected against chance imbalances in covariate distributions and improves credibility overall. We could imagine that even if all things are kept equal, age distribution might be a confounder in randomized clinical trials. 
    \item The standard way of estimating the propensity score is using a logistic regression to estimate probability of treatment. Given what we know about the curse of dimensionality, do you think there might be advantages to using other machine learning algorithms (decision trees, bagging/boosting forests, ensembles, etc.) to estimate propensity scores instead?
    \item \textbf{Your Answer:} Logistic regression performs poorly when you have models that have non-linear relationships. Thus, ensemble methods (e.g., bagging/boosting forests) could provide bias reduction and more consistent confidence interval coverage.It is also the case that models like logistic regression require assumptions that are asy to violation about interactions (between variables) and distributions of variables. Decision trees, for example, seem like  a logical step in estimating propensity scores insofar as, they are able to handle various types of data including categorical, continuous, and ordinal data. Interactions and non-linear relationships are easily modeled also as a result of the natural cleavages that stem from the design of decision trees. 
\end{enumerate}